\chapter*{Acknowledgments}
\markboth{Acknowledgments}{Acknowledgments}
\addcontentsline{toc}{chapter}{Acknowledgments}

This thesis would not have been possible without the contributions, explicit or implicit, of many.
First and foremost, I would like to thank my advisor, Matthias Grossglauser, for his mentorship.
His vision and sharp insights have helped shape much of the research presented in this thesis.
I am also particularly grateful for his never-ending generosity and enthusiasm.

Next, I would like to thank the members of my jury committee:
Martin Jaggi, Michael Gastpar, Balaji Prabhakar, Devavrat Shah and Milan Vojnović.
I thank them for the time they made in their busy schedules to read my dissertation carefully and provide useful comments.
Special thanks go to Patrick Thiran, for taking interest in my work and for his valued feedback.

I have been fortunate to be surrounded by fantastic staff at the lab.
Holly Cogliati-Bauereis provided invaluable help by proof-reading and improving my manuscripts.
Patricia Hjelt, Angela Devenoge and Danielle Alvarez assisted me with the administrative tasks---always with utmost kindness.
Carlos Perez helped me with cutting-edge IT infrastructure needs.
I would like to thank them all.

It was a joy to work with a group of kind-hearted and clever people who gave supportive advice throughout my time at EPFL.
Brunella, many thanks for your friendship and for inspiring me to always give the best of myself.
Young-Jun, I have learned a lot from and with you, thank you for your benevolence.
Victor, thank you for your enthusiasm and your constant positive attitude.
Christina and William, thank you for being the best office mates I could wish for.
Thank you
Farid,
Mohamed,
Julien,
Runwei,
Vincent,
Sébastien,
Farnood,
Daniyar,
Aswin,
Elisa,
Emti,
and Régis!

Finally, I would like to express my gratitude to my family, who supported me every step of the way.
Thank you for your trust in me.
A big heartfelt thank you goes to Ksenia, for making everything in life enjoyable.

\bigskip
 
\noindent\textit{Lausanne, May 15, 2018}
\hfill L.~M.
